Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-07 paper/sec-analysis.tex
more work in the analysis section

Also in the boosted category we require the two
leading subjets of the large-$R$ jet to be relatively
hard, in particular that should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig.~\ref{fig:cutplots22}
we show the distribution in $p_T$ of the leading
and subleading sub-jets in the leading large-$R$ jet in events
of the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background,
motivating our choice of subjet $p_T$ cut.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
 \includegraphics[width=0.48\textwidth]{plots/pt_leadSJ_fj1_noPU.pdf}
 \includegraphics[width=0.48\textwidth]{plots/pt_subleadSJ_fj1_noPU.pdf}
\caption{\small  The $p_T$ distribution of the
  leading (left plot) and
  subleading (right plot) AKT03 subjets within the leading
  Higgs candidate in the boosted category.
  %
  We comparing
  the shapes of the distribution in signal and background events,
  where both
  distributions have been normalized to their total integral.
}
\label{fig:cutplots22}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

In the resolved category,
it is important to understand the $p_T$ distribution
of the four leading small-$R$ jets of the event.
%
As noted in~\cite{deLima:2014dta}
given that in general the boost from the Higgs decays is moderate,
the two subleading jets might be not too hard, and thus it
should be important to ensure that our $p_T^{\rm min}$ cut
is not too strong.
%
This also has important implications for the experimental
trigger requirements in this category.
%
In Fig.~\ref{fig:cutplots23}
we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events.
%
We see that the third and four leading jets are relatively soft.
%
Therefore, we can conclude that....

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
 \includegraphics[width=0.48\textwidth]{plots/pt_smallRjets_res_noPU.pdf}
 \includegraphics[width=0.48\textwidth]{plots/eta_smallRjets_res_noPU.pdf}
 \caption{\small Same as Fig.~\ref{fig:cutplots23}, now for the
   $p_T$ and rapidity distributions of the small-$R$
   jets in the boosted category.
}
\label{fig:cutplots23}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Another important distribution is the rapidity of the large-$R$
(in the boosted category) and the small-$R$ (in the resolved
category) jets.
%
We want to understand by how much the signal efficiency is reduced
by the restriction of the central region, and how would
the signal yield be enhanced if tracking could be extended
to forward rapidities in ATLAS and CMS.
%
Also, it is useful to show the distribution in rapidity of the
individual jets in the resolved category.
%

Next, in Fig.~\ref{fig:mHHinv} we show the invariant mass
of the leading reconstructed Higgs candidates, before the Higgs mass window
cut Eq.~(\ref{higgsmasswindow})
  is applied, for the resolved and boosted categories.
%
The signal distribution is of course peaked at the
nominal Higgs mass of $m_h=125$ GeV.
%
In the boosted category, the background shows no particular
structure, while remarkable in the resolved category the background
also shows a peak-like structure, artificially induced by the
acceptance kinematical cuts on the jets.
%
Note that the applied smearing, which mimics detector resolution effects,
leads to a rather broad distribution which reduces the usefulness
of the Higgs mass window cut.
%
While our Higgs mass window cut is relatively loose,
information of the different shape of the $m_{h}$
distribution will still be exploited by the MVA.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.48\textwidth]{plots/m_H0_res_C1.pdf}
  \includegraphics[width=0.48\textwidth]{plots/m_H0_boost_C1.pdf}
  \caption{\small The invariant mass distribution of the leading
    Higgs candidates on the resolved (left plot) and boosted (right
    plot) categories. 
}
\label{fig:mHHinv}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Another selection requirement for the boosted category is that the two
leading AKT03 subjets of the large-$R$ jet must be relatively hard,
in particular they should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig.~\ref{fig:cutplots22}
we show the distribution in $p_T$ of the leading
and subleading AKT03 subjets in the subleading large-$R$ jet in events
corresponding to the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background.
%
On the other hand, considering the subleading AKT03 subjet, one needs that
the cut in its  $p_T$ should not bee too large, to keep most of the signal events.
%
So as in the case of the $p_T$ cut in the large-$R$ jets, also in this case there
is a compromise between suppressing backgrounds but keeping a large fraction of
signal events.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
 \includegraphics[width=0.48\textwidth]{plots/pt_leadSJ_fj2_noPU.pdf}
 \includegraphics[width=0.48\textwidth]{plots/pt_subleadSJ_fj2_noPU.pdf}
 \caption{\small  Same as Fig.~\ref{fig:cutplots1} for the two leading AKT03
   subjets in the subleading Higgs candidate large-$R$ jet.
}
\label{fig:cutplots22}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Turning to the resolved category, an important aspect to account
in the corresponding selection
cuts is the fact that the $p_T$ distribution
of the four leading small-$R$ jets of the event can be relatively soft,
specially for the subleading jets.
%
As noted in~\cite{deLima:2014dta}, this is a consequence of the fact
that in general the boost from the Higgs decays is moderate.
%
These considerations lead to the requirement that the $p_T$ cut
for the small-$R$ jets cannot be too strong.
%
In Fig.~\ref{fig:cutplots23}
we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events: we observe that both
distributions are peaked at $p_T \le 50$ GeV, with the signal distribution
eventually becoming dominant at large $p_T$.
%
The feasibility to being able to trigger on four small-$R$ jets with a relatively
soft $p_T$ distribution is one of the experimental challenges of
exploiting the resolved category in this final state.
%
In  Fig.~\ref{fig:cutplots23} we also show the
rapidity distribution of the the small-$R$
jets in the resolved category: as well know, the production
is mostly central (more from signal since backgrounds are dominated by
QCD $t$-channel exchange) and thus the standard rapidity cuts
are relatively harmless.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
 \includegraphics[width=0.48\textwidth]{plots/pt_smallRjets_res_noPU.pdf}
 \includegraphics[width=0.48\textwidth]{plots/eta_smallRjets_res_noPU.pdf}
 \caption{\small Same as Fig.~\ref{fig:cutplots1}, now for the
   $p_T$ and rapidity distributions of the small-$R$
   jets corresponding to the resolved selection.
}
\label{fig:cutplots23}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

One of the selection cuts with highest discrimination power is the requirement
that the invariant mass of the Higgs candidate (di)jets must lie within a window
around the nominal Higgs value.
%
In Fig.~\ref{fig:mHHinv} we show the invariant mass
of the leading reconstructed Higgs candidates, before the Higgs mass window
cut Eq.~(\ref{higgsmasswindow})
  is applied, both for the resolved and boosted categories.
%
The signal distribution is of course peaked at the
nominal Higgs mass of $m_h=125$ GeV.
%
In the boosted category, the background shows no particular
structure, while remarkable in the resolved category the background
also shows a peak-like structure, artificially induced by the
acceptance kinematical cuts on the jets.
%
Note that the applied smearing, which mimics detector resolution effects,
leads to a rather broad distribution which reduces the usefulness
of the Higgs mass window cut.
%
While our Higgs mass window cut is relatively loose,
information of the different shape of the $m_{h}$
distribution will still be exploited by the MVA.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.48\textwidth]{plots/m_H0_res_C1d_noPU.pdf}
  \includegraphics[width=0.48\textwidth]{plots/m_H0_bst_C1d_noPU.pdf}
  \caption{\small Same as   Fig.~\ref{fig:cutplots1} for the invariant
    mass distribution of the leading Higgs candidates in the resolved
    (left plot) and boosted (right plot) selections.
}
\label{fig:mHHinv}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
